\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2024

% ready for submission
% \usepackage[final]{neurips_2024}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
\usepackage[preprint]{neurips_style/neurips_2024}

% Remove the conference notice at the bottom of the first page
\makeatletter
\renewcommand{\@notice}{}
\makeatother

% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_2024}

% to avoid loading the natbib package, add option nonatbib:
%    \usepackage[nonatbib]{neurips_2024}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{newunicodechar}  % support for unicode characters
% Define unicode characters that might appear in verbatim
\newunicodechar{θ}{\ensuremath{\theta}}
\newunicodechar{σ}{\ensuremath{\sigma}}
\newunicodechar{μ}{\ensuremath{\mu}}
\newunicodechar{τ}{\ensuremath{\tau}}
\newunicodechar{Δ}{\ensuremath{\Delta}}
\newunicodechar{ε}{\ensuremath{\varepsilon}}
\newunicodechar{γ}{\ensuremath{\gamma}}
\newunicodechar{λ}{\ensuremath{\lambda}}
\newunicodechar{β}{\ensuremath{\beta}}
\newunicodechar{─}{-}
\newunicodechar{├}{+}
\newunicodechar{└}{+}
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage[expansion=false]{microtype}      % microtypography
\usepackage{xcolor}         % colors
\usepackage{amsmath}        % advanced math
\usepackage{amssymb}        % additional math symbols
\usepackage{mathtools}      % more math tools
\usepackage{algorithm}      % algorithm environment
\usepackage{algorithmic}    % algorithmic environment
\usepackage{graphicx}       % include graphics
\usepackage{subcaption}     % subfigures
\usepackage{enumitem}       % better enumerate
\usepackage{bm}             % bold math

\title{SWIVL: Screw-Wrench informed Impedance Variable Learning for bimanual manipulation of an articulated object}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{%
  Kyungseo Park \\
  Student ID : 2020-12295 \\
  Department of Mechanical Engineering \\
  Seoul National University \\
  \texttt{erickun0125@snu.ac.kr} \\
  % examples of more authors
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \AND
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}

\begin{document}

\maketitle

\begin{abstract}
Bimanual manipulation of articulated objects requires simultaneous satisfaction of kinematic constraints and regulation of inter-arm contact forces—capabilities that current imitation-based approaches that lack adaptive force modulation and rely on high-stiffness control. We introduce \textbf{SWIVL} (\textbf{S}crew-\textbf{W}rench informed \textbf{I}mpedance \textbf{V}ariable \textbf{L}earning), a hierarchical control framework bridging cognitive planning with physical execution. SWIVL contributes: (1) twist-driven impedance control via stable imitation vector fields that bypass nonlinear pose-error Jacobians; (2) screw axes based decompositions enabling independent compliance for bulk and internal motions; and (3) wrench-adaptive impedance variable learning via RL that suppresses excessive fighting forces. We evaluate SWIVL on SE(2) planar benchmarks involving revolute rotation and prismatic insertion, demonstrating improved success rates, reduced internal forces compared to imitation learning baselines. 
\end{abstract}

% Main Content
\input{main_contents/introduction}
\input{main_contents/related_work}
\input{main_contents/method}
\input{main_contents/experiments}
\input{main_contents/discussion}

\section{Conclusion}
\label{sec:conclusion}

We introduced \textbf{SWIVL} (Screw-Wrench informed Impedance Variable Learning), a hierarchical control framework that bridges the gap between high-level cognitive planning and low-level physically grounded execution for bimanual manipulation of articulated objects. SWIVL addresses the fundamental limitation of existing learning-based manipulation approaches: their reliance on high-stiffness position control without explicit reasoning about kinematic constraints or inter-arm force interactions.

Our framework contributes three key technical innovations. First, twist-driven impedance control via stable imitation vector fields bypasses the nonlinear pose-error Jacobian inherent in SE(3) impedance formulations by incorporating pose errors directly into reference twists, enabling geometrically consistent compliance under large trajectory deviations. Second, screw axes-based decomposition partitions the twist and wrench spaces into object-centric internal and bulk motion subspaces, allowing independent compliance modulation aligned with the articulated object's kinematic structure. Third, wrench-adaptive impedance variable learning via reinforcement learning enables the policy to discover task-appropriate impedance strategies conditioned on real-time force feedback and object geometry, suppressing harmful fighting forces while maintaining compliant trajectory tracking.

Our SE(2) experiments validate the core hypothesis: explicit encoding of geometric constraints and wrench feedback is essential for robust bimanual manipulation. SWIVL consistently outperforms position control and classical impedance control baselines in success rate while significantly reducing inter-arm fighting forces across revolute and prismatic joint types. These results demonstrate that object-aware motion decomposition and learned impedance modulation provide the physical intelligence that cognitive planners lack.

By operating as a planner-agnostic low-level control layer, SWIVL enables safer deployment of vision-language-action models and other cognitive policies for contact-rich bimanual tasks. The framework's principled integration of screw theory, impedance control, and reinforcement learning offers a foundation for extending physical intelligence to more complex articulated manipulation scenarios in SE(3).
% Bibliography
\bibliographystyle{plainnat}
\bibliography{references}

% Appendix
\appendix

\input{appendix/appendix_notation}
\input{appendix/appendix_impedance_control}
\input{appendix/appendix_se2}
\input{appendix/appendix_learning_settings}

\end{document}
